\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-06-08 paper/ms.tex
More of the ABC model

@@

We show posterior samples in Figure~\ref{fig:abc_results}.

@@

%\begin{table}[ht!]
%\centering
%\small
%\begin{tabular}{|l|l|l|}
%\hline
%Bayesian computation		&		Statistical mechanics		&		ABC\\
%\hline
%Parameter space	$\Theta$	&		Phase space	$\Omega$ 			& Product space $\Theta \times \mathcal{D}$\\
%Posterior $p(\params | D,I)$   &  Canonical distribution $p(\boldsymbol{x}; T)$     & \\ 
%Marginal likelihood $P(D|I)$	&	Partition function $Z(T)$	&   Marginal likelihood $p(D\in | I)$\\
%\hline
%\end{tabular}
%\caption{\it The relationship between standard Bayesian computation, statistical
%mechanics, and ABC. In each case Monte Carlo methods are used to calculate
%integrals over a space. In standard Bayesian computation, it is the parameter
%space of a model that is usually of interest, whereas in ABC it is the space
%of possible parameter values {\it and} data sets.
%\label{tab:relation}}
%\end{table}

%\begin{figure}[ht!]
%\centering
%\includegraphics[scale=0.5]{figures/joint.pdf}
%\caption{\it The joint prior for parameters $\params$ and data $\data$.
%\label{fig:joint}}
%\end{figure}

%\section{Parameterising the data space}
%In standard MCMC-based ABC, proposal moves are made by first proposing a
%change to the parameters (from $\params$ to $\params'$), and then generating
%a mock dataset $\data'$ from $p(\data | \params')$. As a Metropolis proposal
%for exploring the product space $\Theta \times \mathcal{D}$ this is a very
%aggressive choice, since the proposed value of the
%dataset $\params'$ is likely to be very different from the previous one.
%One of the most elementary pieces of advice we receive when learning the
%Metropolis algorithm is to be careful about how ambitious our proposal
%distributions are, since tiny and huge proposal moves are both very
%inefficient. The performance of MCMC-based ABC can therefore be improved by
%alternative choices of proposal distribution.

%In some cases, we might be able to find efficient proposal distributions for exploring $\Theta \times \mathcal{D}$ by having insight into the particular
%problem we're working on. However, there are also techniques that are
%completely general in that they should work on any ABC-type problem.

%Consider the process of generating a mock dataset. Whatever else is involved,
%this process will involve calls to an underlying random number generator.
%Perhaps a function {\tt rand()}, generating variables from a Uniform(0, 1)
%distribution, is called $n$ times in the course of generating a dataset.
%Usually, the simulation process has the property that, if $\theta'$ is
%close to $\theta$, then the same sequence of random numbers returned by
%{\tt rand()} would produce a dataset $\data'$ that is very similar to $\data$.
%This suggests an alternative parameterisation of
%the product space $\Theta\times\mathcal{D}$: instead of using the one that
%naturally describes the data, we can use a coordinate system defined by
%$\theta$ together with $\{u_1, ..., u_n\}$, the $n$ uniform random numbers
%used to construct a mock dataset.

%In this alternative coordinate system, a proposed move
%to change $\params$ while keeping
%$\{u_i\}$ fixed results in a proposed new dataset that is similar to the
%previous one, assuming the change to $\params$ is small. The standard procedure
%of generating an entirely new mock dataset each step is now viewed as a
%proposal where $\params$ is moved {\it and all of the $u_i$ coordinates are
%re-generated from their iid uniform prior}.

@@

\section{Specifying models in Python with dnest4.builder}

>>>>

\code{RJObject}'s \code{print} function prints the dimensionality of each
component and the maximum value of $N$, followed by the hyperparameters,
then parameter zero of each component (zero padded when $N<N_{\rm max}$),
parameter one for each component, and so on.

@@

The proposal involves changing either \code{mu}, \code{log\_sigma}, or
one of the \code{n}s:
\begin{CodeChunk}
\begin{CodeInput}
double MyModel::perturb(RNG& rng)
{
	int which = rng.rand_int(3);

	if(which == 0)
	{
		mu += 20.*rng.randh();
		wrap(mu, -10., 10.);
	}
	if(which == 1)
	{
		log_sigma += 20.*rng.randh();
		wrap(log_sigma, -10., 10.);
	}
	if(which == 2)
	{
		int i = rng.rand_int(n.size());
		n[i] = rng.randn();
	}

	return 0.0;
}
\end{CodeInput}
\end{CodeChunk}

The ``log-likelihood'' is really minus the discrepancy function.
Since we parameterized joint (parameter, data) space using the
$n$ variables instead of the data $\{x_i\}$ itself, we must
``assemble'' the simulated dataset in order to evaluate the
discrepancy function:
\begin{CodeChunk}
\begin{CodeInput}
double MyModel::log_likelihood() const
{
    double x_min = Data::get_instance().get_x_min();
    double x_max = Data::get_instance().get_x_max();

    double sigma = exp(log_sigma);

    // Assemble fake dataset
    vector<double> x_fake = n;
    for(size_t i=0; i<x_fake.size(); i++)
        x_fake[i] = mu + sigma*x_fake[i];

    // Goodness
    double logL = 0.;
    logL -= pow(*min_element(x_fake.begin(), x_fake.end()) - x_min, 2);
    logL -= pow(*max_element(x_fake.begin(), x_fake.end()) - x_max, 2);

    return logL;
}
\end{CodeInput}
\end{CodeChunk}

\subsection{ABC Example Results}
We show posterior samples in Figure~\ref{fig:abc_results}.

@@

@@

%\section{Specifying models in Python with dnest4.builder}
